---
title: "RS sampling of Aveiro Lagoon"
editor_options: 
  chunk_output_type: console
---

<!-- ::: {style="text-align: center; margin-top: 20px;"} -->

<!-- <button onclick="window.open(&#39;https://github.com/SigOiry/Coring_Carbon_Seagrasses/raw/refs/heads/main/Data/QGIS/Qgis_Project_Cadiz.zip&#39;, &#39;_blank&#39;);" style="padding: 10px 20px; background-color: #4CAF50; color: white; border: none; border-radius: 5px; cursor: pointer;"> -->

<!-- Download the Qgis project of Cadiz Bay -->

<!-- </button> -->

<!-- ::: -->

```{r update gitignore}
#| cache: false
#| echo: false
#| warning: false
#| eval: false

library(fs)

# Define the size threshold in bytes (100 MB)
size_threshold <- 100 * 1024 * 1024

# Get a list of all files in the repository
files <- dir_ls(recurse = TRUE, type = "file")

# Filter files larger than the size threshold
large_files <- files[file_info(files)$size > size_threshold]

# Check if there are large files
if (length(large_files) > 0) {
  # Read the existing .gitignore file if it exists
  gitignore_path <- ".gitignore"
  gitignore_content <- if (file_exists(gitignore_path)) {
    readLines(gitignore_path)
  } else {
    character(0)
  }
  
  # Identify files not already in .gitignore
  files_to_add <- large_files[!large_files %in% gitignore_content]
  
  # Append new large files to .gitignore
  if (length(files_to_add) > 0) {
    writeLines(c(gitignore_content, files_to_add), gitignore_path)
    message(length(files_to_add), " file(s) added to .gitignore.")
  } else {
    message("No new files to add to .gitignore.")
  }
} else {
  message("No files larger than 100 MB found.")
}
```

::: {style="text-align: center; margin-top: 20px;"}
<button onclick="window.open(&#39;https://github.com/SigOiry/Coring_Carbon_Seagrasses/raw/refs/heads/Aveiro_Coastal_Lagoon/Output/Aveiro/Macrofauna_Grid_Aveiro.zip&#39;, &#39;_blank&#39;);" style="padding: 10px 20px; background-color: #4CAF50; color: white; border: none; border-radius: 5px; cursor: pointer;">

Download extracted data from the Macrofauna Grid
</button>
:::

# Methods

Macrofauna sampling was conducted during the field campaign in the Ria de Aveiro Lagoon, which took place from July 14th to 18th, 2025. Sampling effort was distributed across two sites: “Rewilded” and “Natural.” In this study, the sampling points will be compared with metrics derived from Sentinel-2 imagery. The Sentinel-2 image used was acquired on July 15th, 2025, and corresponds to a Level-2A product, which has been atmospherically corrected and orthorectified by the data provider using Sen2Cor.

```{r leafletmap}
#| fig-cap: Map of macrofauna sampling points of the Rewrite Project, in Ria de Aveiro Lagoon. 
#| echo: false
#| error: false
#| message: false
#| warning: false
#| out-width: "95%"

library(leaflet)
library(htmlwidgets)
library(jsonlite)
library(sf)
library(tidyverse)

sampling_point <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp")

bb <- st_bbox(sampling_point) %>% 
  as.numeric()

bb_poly_rewilded <- sampling_point %>%          # keep it an sf
  filter(site == "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()                                   # → sfc_POLYGON

bb_poly_natural <- sampling_point %>% 
  filter(site != "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()

bbox_sf <- rbind(
  st_sf(site = "rewilded", geometry = bb_poly_rewilded),
  st_sf(site = "natural",  geometry = bb_poly_natural)
)

  
pal <- colorFactor(
  palette = "Set1",          
  domain  = sampling_point$site
)

m <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  fitBounds(
    lng1 = bb[1], lat1 = bb[2],
    lng2 = bb[3], lat2 = bb[4]
  )%>%                                  
  addCircleMarkers(
    data        = sampling_point,            
    radius      = 6,
    stroke      = FALSE,
    fillColor   = ~pal(site),               
    fillOpacity = 0.9,
    label       = ~name,                     
    labelOptions = labelOptions(
      direction = "auto",
      textsize  = "12px",
      style     = list("font-weight" = "normal")
    )
  ) %>%
  addLegend(                                
    "bottomright",
    pal     = pal,
    values  = sampling_point$site,
    title   = "Site",
    opacity = 1
  ) %>% 
  leafem::addMouseCoordinates(css = list(
    "font-size" = "20px",
    "text-align" = "center",
    "background-color" = "white",
    "color" = "rgb(0, 0, 0)"
  )) %>% 
  
  # (6) Measurement tool (click to start drawing; click again to add points; double‐click to finish)
  addMeasure(
    position           = "topright",     # control’s corner on the map
    primaryLengthUnit  = "meters",       # or "kilometers", "miles", etc.
    primaryAreaUnit    = "sqmeters",     # or "hectares", "acres", etc.
    activeColor        = "#3D535D",      # line color while drawing
    completedColor     = "#7D4479",      # line color after finishing
    captureZ           = FALSE           # if your CRS has elevation you can include Z
  ) 

m 

```

## Images processing

All image processing was performed in R, using the *terra*, *sf* and the *tidyverse* packages.

### NDVI

The NDVI as been computed using the following equation:

$$
NDVI = (B8-B4)/(B8+B4)
$$

### Seagrass Percent Cover

The methodology applied in this study is adapted from [Zoffoli et al. (2021)](https://doi.org/10.1016/j.ecolind.2021.108033).

Seagrass Percentage Cover (SPC) was derived from the NDVI using the equation established by [Zoffoli et al. (2020)](https://doi.org/10.1016/j.rse.2020.112020):

$$
SPC = 172.06 * NDVI - 22.18
$$ \### SHSI

This methodology is currently under review but a draft of the submitted paper can be found [here](https://sigoiry.github.io/HeatWave_Seagrasses).

Based on the observed spectral changes in seagrasses exposed to HWs, we developed a new radiometric index to better detect the radiometric change caused by the HW. The browning of the leaves was characterized by substantial radiometric changes in both the green and red-edge spectral regions. The Seagrass Heat Shock Index (SHSI) was introduced as the reflectance line height at 740 nm, compared to the 560 - 842 nm baseline. Namely the SHSI subtracts the reflectance observed at 740 nm to the linearly interpolated reflectance between 560 and 842 nm, so that the index is positive in the case of brown, HW-impacted seagrass leaves, and negative in the case of green, non-impacted leaves. Compared to [the submited version of the index](https://sigoiry.github.io/HeatWave_Seagrasses/#optical-measurements), the SHSI used in this study has been normalized to constrain its values strictly between -1 and 1:

$$
\text{SHSI} = (I_{SHSI} - R(740))/(I_{SHSI} + R(740))
$$

where :

$$
I_{SHSI} = R(560) + \tau    [R(842) - R(560)]
$$ and :

$$
\tau = \frac{740 - 560}{842 - 560}
$$

where $R(560)$, $R(740)$, and $R(842)$ represent the reflectance at 560, 740, and 842 nm, respectively, and $\tau$ is equal to a contant of 0.64. These wavelengths were selected to align with the spectral resolution of satellites missions such as Sentinel-2, for broader remote sensing application.

## Results

```{r overall map}
#| echo: false
#| eval: false
#| warning: false


library(MapRs)
library(terra)

sampling_point <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp")

bb <- st_bbox(sampling_point) %>% 
  as.numeric()

bb_poly_rewilded <- sampling_point %>%          # keep it an sf
  filter(site == "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()                                   # → sfc_POLYGON

bb_poly_natural <- sampling_point %>% 
  filter(site != "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()

bbox_sf <- rbind(
  st_sf(site = "rewilded", geometry = bb_poly_rewilded),
  st_sf(site = "natural",  geometry = bb_poly_natural)
) %>% 
  dplyr::filter(site == "rewilded")

write_sf(bbox_sf, "ShinyApp/Bourgneuf_Site_Selection/Meadow_samplingSite.shp")

S2_img <- MapRs::Read_S2("Data/SAFE/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315.SAFE") 

bbox_32629 <- st_transform(bbox_sf, st_crs(S2_img)) %>% 
  vect() %>% 
  buffer(20)




```

```{r rewilded processing}
#| echo: false
#| eval: false
#| warning: false

library(MapRs)

sampling_point <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp")

sampling_point_32629 <- sampling_point %>% 
  st_transform(st_crs(S2_img)) 

bb <- st_bbox(sampling_point) %>% 
  as.numeric()

bb_poly_rewilded <- sampling_point %>%          # keep it an sf
  filter(site == "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()                                   # → sfc_POLYGON

bb_poly_natural <- sampling_point %>% 
  filter(site != "rewilded") %>% 
  st_bbox() %>% 
  st_as_sfc()

bbox_sf <- rbind(
  st_sf(site = "rewilded", geometry = bb_poly_rewilded),
  st_sf(site = "natural",  geometry = bb_poly_natural)
)
  # dplyr::filter(site == "rewilded")

write_sf(bbox_sf, "ShinyApp/Bourgneuf_Site_Selection/Meadow_samplingSite.shp")

S2_img <- MapRs::Read_S2("Data/SAFE/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315.SAFE") 

bbox_32629 <- st_transform(bbox_sf, st_crs(S2_img)) %>% 
  vect() %>% 
  buffer(20)

mask <- vect("Data/Aveiro_Lagoon/SHP/Intertidal_masks_Aveiro.shp")

S2_img_cropped <- (S2_img) %>% 
  crop(mask, mask = T) - 1000 ### Crop and Correct the offset applied on the baseline after 4.00



NDVI <- (S2_img_cropped$B08-S2_img_cropped$B04)/(S2_img_cropped$B08+S2_img_cropped$B04)

SPC <- 172.06*NDVI-22.18

values(SPC)[values(SPC) > 100] <- 100
values(SPC)[values(SPC) < 0] <- 0



mn  <- app(S2_img_cropped, min, na.rm = TRUE)   # one‑layer SpatRaster of cellwise minima
mx  <- app(S2_img_cropped, max, na.rm = TRUE)   # one‑layer SpatRaster of cellwise maxima
rng <- mx - mn

# Scale to [0, 1]
S2_img_cropped_STD <- (S2_img_cropped - mn) / rng

# Guard against pixels where all band values are equal or all NA (rng == 0 or NA)
S2_img_cropped_STD <- ifel(rng == 0 | is.na(rng), NA, x01)  # or replace NA with 0 if you prefer

# (optional) keep original layer names
names(S2_img_cropped_STD) <- names(S2_img_cropped)

SHSI <- ((S2_img_cropped_STD$B03+0.64*(S2_img_cropped_STD$B08-S2_img_cropped_STD$B03))-S2_img_cropped_STD$B06)/((S2_img_cropped_STD$B03+0.64*(S2_img_cropped_STD$B08-S2_img_cropped_STD$B03))+S2_img_cropped_STD$B06)

SHSI_V2 <- ((S2_img_cropped_STD$B02+0.708*(S2_img_cropped_STD$B08-S2_img_cropped_STD$B02))-S2_img_cropped_STD$B06)/((S2_img_cropped_STD$B02+0.708*(S2_img_cropped_STD$B08-S2_img_cropped_STD$B02))+S2_img_cropped_STD$B06)


stk_metrics <- c(SPC,NDVI,SHSI_V2)
names(stk_metrics) <- c("SPC","NDVI","SHSI")

writeRaster(stk_metrics, "Data/Aveiro_Lagoon/Output/metrics_all.tif", overwrite=T)
```

```{r stat maps}
#| echo: false
#| eval: true
#| warning: false

library(terra)
library(tidyverse)

mask <- vect("Data/Aveiro_Lagoon/SHP/mask_rewilded.shp")
rast <- rast("Data/Aveiro_Lagoon/Output/metrics_all.tif") %>% 
  crop(mask, mask = T) %>% 
  as.data.frame()

mean_NDVI <- round(mean(rast$NDVI),2)
max_NDVI <- round(max(rast$NDVI),2)

mean_SPC <- round(mean(rast$SPC),2)
max_SPC <- round(max(rast$SPC),2)

mean_SHSI <- round(mean(rast$SHSI),2)
max_SHSI <- round(max(rast$SHSI),2)

mask_nat <- vect("Data/Aveiro_Lagoon/SHP/mask_natural.shp")
rast_nat <- rast("Data/Aveiro_Lagoon/Output/metrics_all.tif") %>% 
  crop(mask_nat, mask = T) %>% 
  as.data.frame()

mean_NDVI_nat <- round(mean(rast_nat$NDVI),2)
max_NDVI_nat <- round(max(rast_nat$NDVI),2)

mean_SPC_nat <- round(mean(rast_nat$SPC),2)
max_SPC_nat <- round(max(rast_nat$SPC),2)

mean_SHSI_nat <- round(mean(rast_nat$SHSI),2)
max_SHSI_nat <- round(max(rast_nat$SHSI),2)

```

@fig-Maps_all shows the spatial distribution of the three satellite-derived metrics considered for the rewilded study site in the Ria de Aveiro Lagoon. The average NDVI within the extent of the macrofauna sampling stations was only `r mean_NDVI`, with a maximum of `r max_NDVI` observed in the central part of the meadow. When converted to Seagrass Percent Cover (SPC), this site exhibited a low average SPC of `r mean_SPC`%, just above the 20% threshold commonly used to define a seagrass meadow ([Zoffoli et al., 2020](https://www.sciencedirect.com/science/article/pii/S0034425720303904#s0070)). In the densest part of the meadow, SPC reached up to `r max_SPC`%. 
Regarding the SHSI, which serves as an indicator of heat-induced color change in seagrass, the average value was `r mean_SHSI`, suggesting that the meadow was not affected by a heatwave at the time of observation.

@fig-Maps_all shows the spatial distribution of the three satellite-derived metrics considered for the natural study site in the Ria de Aveiro Lagoon. The average NDVI within the extent of the macrofauna sampling stations was `r mean_NDVI_nat`, with a maximum of `r max_NDVI_nat` observed in the central part of the meadow. When converted to Seagrass Percent Cover (SPC), this site exhibited a higher average SPC compared to the rewilded site, with an average of `r mean_SPC_nat`%. In the densest part of the meadow, SPC reached up to `r max_SPC_nat`%. 
Regarding the SHSI, which serves as an indicator of heat-induced color change in seagrass, the average value was `r mean_SHSI_nat`, suggesting that the meadow was not affected by a heatwave at the time of observation. However, some pixel of that study sites are showing SHSI values close to 0 and sometime positive, maybe suggesting heat-induce discoloration over some part of the meadow.

```{r Maps Metrics}
#| fig-cap: Map of satellite derived metrics in the study sites of Ria de Aveiro Lagoon
#| label: fig-Maps_all
#| echo: false
#| error: false
#| message: false
#| warning: false
#| out-width: "100%"

library(leafsync)
library(terra)
library(tidyverse)
library(sf)



####################################################################### NDVI
pal_NDVI <- colorNumeric(
  palette = RColorBrewer::brewer.pal(11, "Spectral"),
  domain  = NULL,
  na.color = "transparent",
  reverse = T
)

mask <- vect("Data/Aveiro_Lagoon/SHP/Intertidal_masks_Aveiro.shp") %>% 
  project(crs("EPSG:4326"))

stk_metrics <- rast("Data/Aveiro_Lagoon/Output/metrics_all.tif") %>% 
  project(crs("EPSG:4326")) %>% 
  crop(mask, mask = T)

sampling_point <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp")
  # st_transform(crs("EPSG:32629"))

sampling_polygons <- vect("ShinyApp/Bourgneuf_Site_Selection/Meadow_samplingSite.shp")

pal <- colorFactor(
  palette = "Set1",          
  domain  = sampling_point$site
)

# extracted_values <- terra::extract(stk_metrics,sampling_point)

values(stk_metrics)[values(stk_metrics$NDVI) <0] <- NA

NDVI <- stk_metrics$NDVI

# build the leaflet map
NDVI_map <- leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  # add the raster, assign a layerId so leafem knows which layer to query
  addRasterImage(
    x       = NDVI, 
    # project = TRUE,
    colors  = pal_NDVI, 
    opacity = 1, 
    layerId = "NDVI",
    group = "NDVI"
  ) %>% 
  # add a continuous legend
  addLegend(
    pal    = pal_NDVI, 
    values = values(NDVI),
    title  = "NDVI value",
    position = "bottomright"
  ) %>% 
  # wire up click‐popup: show value with 2 decimals
  leafem::addImageQuery(
    # map    = .,
    x = NDVI,
    project = TRUE,
    layerId= "NDVI",
    type   = "mousemove",
    digits = 2,
    prefix = "",
    position = "bottomright",

  ) %>%                                  
  addCircleMarkers(
    data        = sampling_point,            
    radius      = 6,
    stroke      = FALSE,
    fillColor   = ~pal(site),               
    fillOpacity = 0.9,
    label       = ~name,                     
    labelOptions = labelOptions(
      direction = "auto",
      textsize  = "12px",
      style     = list("font-weight" = "normal")
    )
  ) %>%
  
  leafem::addMouseCoordinates(css = list(
    "font-size" = "20px",
    "text-align" = "center",
    "background-color" = "white",
    "color" = "rgb(0, 0, 0)"
  )) %>%
  
  # (6) Measurement tool (click to start drawing; click again to add points; double‐click to finish)
  addMeasure(
    position           = "topright",     # control’s corner on the map
    primaryLengthUnit  = "meters",       # or "kilometers", "miles", etc.
    primaryAreaUnit    = "sqmeters",     # or "hectares", "acres", etc.
    activeColor        = "#3D535D",      # line color while drawing
    completedColor     = "#7D4479",      # line color after finishing
    captureZ           = FALSE           # if your CRS has elevation you can include Z
  ) %>% 
      addLayersControl(overlayGroups = "NDVI")



##################################################### SPC 

# pal_SPC <- colorNumeric(
#   palette = RColorBrewer::brewer.pal(11, "Spectral"),
#   domain  = NULL,
#   na.color = "transparent",
#   reverse = T
# )

pal_SPC <- colorNumeric(
  palette  = colorRampPalette( rev(RColorBrewer::brewer.pal(11, "RdYlGn")) )(255),
  domain   = c(20, 80),      # adapt if your data range is different
  na.color = "transparent"
)

pal_SPC(80)

SPC <- stk_metrics$SPC

values(SPC)[values(SPC) <20] <- NA
values(SPC)[values(NDVI) <0] <- NA

ticks <- seq(20, 80, by = 10)    # 0, 20, 40, 60, 80, 100

pal_SPC_clamped <- function(z) pal_SPC(pmin(pmax(z, 20), 80))

# build the leaflet map
SPC_map <- leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  # add the raster, assign a layerId so leafem knows which layer to query
  addRasterImage(
    x       = SPC, 
    # project = TRUE,
    colors  = pal_SPC_clamped,
    opacity = 1, 
    layerId = "SPC",
    group = "SPC",
  ) %>% 
  # add a continuous legend
  addLegend(
    pal       = pal_SPC,
    values    = ticks,           # fewer values → shorter legend
    title     = "SPC (%)",
    position  = "bottomright",
    labFormat = labelFormat(suffix = "%")   # optional nice labels
  ) %>% 
  # wire up click‐popup: show value with 2 decimals
  leafem::addImageQuery(
    # map    = .,
    x = SPC,
    project = TRUE,
    layerId= "SPC",
    type   = "mousemove",
    digits = 2,
    prefix = "",
    position = "bottomright",

  ) %>%
  
  leafem::addMouseCoordinates(css = list(
    "font-size" = "20px",
    "text-align" = "center",
    "background-color" = "white",
    "color" = "rgb(0, 0, 0)"
  )) %>%                                  
  addCircleMarkers(
    data        = sampling_point,            
    radius      = 6,
    stroke      = FALSE,
    fillColor   = ~pal(site),               
    fillOpacity = 0.9,
    label       = ~name,                     
    labelOptions = labelOptions(
      direction = "auto",
      textsize  = "12px",
      style     = list("font-weight" = "normal")
    )
  ) %>%
  
  # (6) Measurement tool (click to start drawing; click again to add points; double‐click to finish)
  addMeasure(
    position           = "topright",     # control’s corner on the map
    primaryLengthUnit  = "meters",       # or "kilometers", "miles", etc.
    primaryAreaUnit    = "sqmeters",     # or "hectares", "acres", etc.
    activeColor        = "#3D535D",      # line color while drawing
    completedColor     = "#7D4479",      # line color after finishing
    captureZ           = FALSE           # if your CRS has elevation you can include Z
  ) %>% 
      addLayersControl(overlayGroups = "SPC")

#################SPC##################################################### SPC 

pal_SHSI <- colorNumeric(
  palette  = colorRampPalette(c("#228B22",   # −0.1 → green
                                "white",    #  0   → white
                                "#8B4513")) #  0.1 → brown
              (255),           # 255‑step gradient
  domain   = c(-0.1, 0.1),
  na.color = "transparent"
)

pal_SHSI_clamped <- function(z) pal_SHSI(pmax(pmin(z, 0.1), -0.1))

SHSI <- stk_metrics$SHSI

NDVI_mask <- stk_metrics$NDVI

# values(SHSI)[values(NDVI_mask) <0] <- NA
# values(SHSI)[values(SHSI) < -0.1] <- -0.1
# values(SHSI)[values(SHSI) > 0.1] <- 0.1

# build the leaflet map
SHSI_map <- leaflet() %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  # add the raster, assign a layerId so leafem knows which layer to query
  addRasterImage(
    x       = SHSI, 
    # project = TRUE,
    colors  = pal_SHSI_clamped , 
    opacity = 1, 
    layerId = "SHSI",
    group = "SHSI"
  ) %>% 
  # add a continuous legend
  addLegend(
    pal    = pal_SHSI, 
    values = c(-0.1,0.1),
    title  = "SHSI value",
    position = "bottomright",
    ) %>% 
  # wire up click‐popup: show value with 2 decimals
  leafem::addImageQuery(
    # map    = .,
    x = SHSI,
    project = TRUE,
    layerId= "SHSI",
    type   = "mousemove",
    digits = 2,
    prefix = "",
    position = "bottomright",

  ) %>%                                  
  addCircleMarkers(
    data        = sampling_point,            
    radius      = 6,
    stroke      = FALSE,
    fillColor   = ~pal(site),               
    fillOpacity = 0.9,
    label       = ~name,                     
    labelOptions = labelOptions(
      direction = "auto",
      textsize  = "12px",
      style     = list("font-weight" = "normal")
    )
  ) %>%
  
  leafem::addMouseCoordinates(css = list(
    "font-size" = "20px",
    "text-align" = "center",
    "background-color" = "white",
    "color" = "rgb(0, 0, 0)"
  )) %>%
  
  # (6) Measurement tool (click to start drawing; click again to add points; double‐click to finish)
  addMeasure(
    position           = "topright",     # control’s corner on the map
    primaryLengthUnit  = "meters",       # or "kilometers", "miles", etc.
    primaryAreaUnit    = "sqmeters",     # or "hectares", "acres", etc.
    activeColor        = "#3D535D",      # line color while drawing
    completedColor     = "#7D4479",      # line color after finishing
    captureZ           = FALSE           # if your CRS has elevation you can include Z
  ) %>% 
      addLayersControl(overlayGroups = "SHSI")


sync(NDVI_map, SPC_map, SHSI_map, sync = "all",ncol = 2)
```

```{r point exctractions}
#| echo: false
#| eval: false
#| warning: false

sampling_point_rewilded <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp") %>% 
  st_transform(crs("EPSG:32629")) %>%
  dplyr::filter(site == "rewilded") %>% 
  vect()

stk_rewilded <- rast("Data/Aveiro_Lagoon/Output/metrics_all.tif") %>% 
  crop(sampling_point_rewilded)

df_rewilded <- terra::extract(stk_rewilded,sampling_point_rewilded)

df_points_rewilded <- sampling_point_rewilded %>% 
  as_tibble() %>% 
  mutate(ID = 1:nrow(.)) %>% 
  left_join(df_rewilded, by = "ID") %>% 
  dplyr::select(-ID)

sampling_point_natural <- read_sf("Data/Aveiro_Lagoon/SHP/Aveiro grid/Aveiro_grid_info.shp") %>% 
  st_transform(crs("EPSG:32629")) %>%
  dplyr::filter(site == "natural") %>% 
  vect()

stk_natural <- rast("Data/Aveiro_Lagoon/Output/metrics_all.tif")%>% 
  crop(sampling_point_natural)

df_natural<- terra::extract(stk_natural,sampling_point_natural)

df_points_natural <- sampling_point_natural %>% 
  as_tibble() %>% 
  mutate(ID = 1:nrow(.)) %>% 
  left_join(df_natural, by = "ID") %>% 
  dplyr::select(-ID)

df_out <- df_points_natural %>% 
  rbind(df_points_rewilded)

write.csv(df_out, "Output/Aveiro/Aveiro_MacrofaunaGrid_Satellite_Metrics/df_sampling_points_Aveiro.csv", row.names = F)
writeRaster(stk_rewilded,"Output/Aveiro/Aveiro_MacrofaunaGrid_Satellite_Metrics/Rewilded_Metrics.tiff", overwrite = T)
writeRaster(stk_natural,"Output/Aveiro/Aveiro_MacrofaunaGrid_Satellite_Metrics/Natural_Metrics.tiff", overwrite = T)

```

```{r zip outputs}
#| echo: false
#| message: false
#| warning: false
#| cache: false
#| eval: false


# ---- settings ----
folder_to_zip <- "Output/Aveiro/Aveiro_MacrofaunaGrid_Satellite_Metrics"   # directory whose contents you want to archive
zip_file      <- "Output/Aveiro/Macrofauna_Grid_Aveiro.zip"

# ---- create the zip ----
# list all files (recursive = TRUE grabs sub-folders too; set FALSE if you only
# want the top-level files)
files <- list.files(folder_to_zip, recursive = TRUE, full.names = TRUE)

# make sure you’re in the folder so the archive has relative paths, not full paths

utils::zip(zipfile = zip_file, files = files)
```


